---
title: "Multiassay Experiments"
author: "Jeffrey Long"
date: "8/31/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("MultiAssayExperiment")

library(MultiAssayExperiment)
library(GenomicRanges)
library(SummarizedExperiment)
# install.packages("RaggedExperiment")   ## Not available for R v.3.5.1, v.3.5.3, or v.3.6.0
# library(RaggedExperiment) 
```


## Vignettes

[MAE Cheatsheet](http://127.0.0.1:20915/library/MultiAssayExperiment/doc/MultiAssayExperiment_cheatsheet.pdf)

[MultiAssay Experiment- The Integrative Bioconductor Container](http://127.0.0.1:20915/library/MultiAssayExperiment/doc/MultiAssayExperiment.html)



```{r}
browseVignettes("MultiAssayExperiment")


```


### MAE- The Integrative Bioconductor Container


```{r}
empty <- MultiAssayExperiment()
empty
slotNames(empty)
class(experiments(empty)) # ExperimentList

# Primary Data- colData
patient.data <- data.frame(sex=c("M", "F", "M", "F"),
    age=38:41,
    row.names=c("Jack", "Jill", "Bob", "Barbara"))
patient.data
class(sampleMap(empty)) # DataFrame

# Instances of sampleMap without colData
exprss1 <- matrix(rnorm(16), ncol = 4,
        dimnames = list(sprintf("ENST00000%i", sample(288754:290000, 4)),
                c("Jack", "Jill", "Bob", "Bobby")))
exprss2 <- matrix(rnorm(12), ncol = 3,
        dimnames = list(sprintf("ENST00000%i", sample(288754:290000, 4)),
                c("Jack", "Jane", "Bob")))
doubleExp <- list("methyl 2k"  = exprss1, "methyl 3k" = exprss2)
simpleMultiAssay <- MultiAssayExperiment(experiments=doubleExp)
simpleMultiAssay
colData(simpleMultiAssay)
# Providing colData
simpleMultiAssay2 <- MultiAssayExperiment(experiments=doubleExp, colData=patient.data)
simpleMultiAssay2
colData(simpleMultiAssay2)
```

### Key Points

* one row of $colData$ can map to zero, one, or more columns in any $ExperimentList$ element

* each row of $colData$ must map to at least one column in at least one $ExperimentList$ element

* each column of each $ExperimentList$ element must map to exactly one row of $colData$.

These relationships are defined by the $sampleMap$.

```{r}
# Creating an MAE

# Expression data
(arraydat <- matrix(seq(101, 108), ncol=4,
    dimnames=list(c("ENST00000294241", "ENST00000355076"),
    c("array1", "array2", "array3", "array4"))))
coldat <- data.frame(slope53=rnorm(4),
    row.names=c("array1", "array2", "array3", "array4"))

exprdat <- SummarizedExperiment(arraydat, colData=coldat)
exprdat
# Map colData to exprdata sample names
(exprmap <- data.frame(primary=rownames(patient.data)[c(1, 2, 4, 3)],
                       colname=c("array1", "array2", "array3", "array4"),
                       stringsAsFactors = FALSE))

# Methylation data as a matrix
(methyldat <-
   matrix(1:10, ncol=5,
          dimnames=list(c("ENST00000355076", "ENST00000383706"),
                        c("methyl1", "methyl2", "methyl3",
                          "methyl4", "methyl5"))))
# Map colData to methyldat
(methylmap <- data.frame(primary = c("Jack", "Jack", "Jill", "Barbara", "Bob"),
    colname = c("methyl1", "methyl2", "methyl3", "methyl4", "methyl5"),
    stringsAsFactors = FALSE))

# microRNA data
(microdat <- matrix(201:212, ncol=3,
                    dimnames=list(c("hsa-miR-21", "hsa-miR-191",
                                    "hsa-miR-148a", "hsa-miR148b"),
                                  c("micro1", "micro2", "micro3"))))
# Map colData sample names to microdat sample names
(micromap <- data.frame(primary = c("Jack", "Barbara", "Bob"),
    colname = c("micro1", "micro2", "micro3"), stringsAsFactors = FALSE))

# Finally create a dataset of class RangedSummarizedExperiment
nrows <- 5; ncols <- 4
counts <- matrix(runif(nrows * ncols, 1, 1e4), nrows)
rowRanges <- GRanges(rep(c("chr1", "chr2"), c(2, nrows - 2)),
    IRanges(floor(runif(nrows, 1e5, 1e6)), width=100),
    strand=sample(c("+", "-"), nrows, TRUE),
    feature_id=sprintf("ID\\%03d", 1:nrows))
names(rowRanges) <- letters[1:5]
colData <- DataFrame(Treatment=rep(c("ChIP", "Input"), 2),
    row.names= c("mysnparray1", "mysnparray2", "mysnparray3", "mysnparray4"))
rse <- SummarizedExperiment(assays=SimpleList(counts=counts),
    rowRanges=rowRanges, colData=colData)
# Map colData samples to RangedSummarizedExperiment
(rangemap <-
    data.frame(primary = c("Jack", "Jill", "Bob", "Barbara"),
    colname = c("mysnparray1", "mysnparray2", "mysnparray3", "mysnparray4"),
        stringsAsFactors = FALSE))

# sampleMap Creation
listmap <- list(exprmap, methylmap, micromap, rangemap)
names(listmap) <- c("Affy", "Methyl 450k", "Mirna", "CNV gistic")
listmap
# Convenience function listToMap to convert to dataframe
dfmap <- listToMap(listmap)
dfmap
# Can revert dfmap to list
mapToList(dfmap, "assay")
#########################################
# Experimental data as a list()
objlist <- list("Affy" = exprdat, "Methyl 450k" = methyldat,
    "Mirna" = microdat, "CNV gistic" = rse)
# Creation of the MAE class object
myMultiAssay <- MultiAssayExperiment(objlist, patient.data, dfmap)
myMultiAssay
experiments(myMultiAssay)
colData(myMultiAssay)
sampleMap(myMultiAssay)
metadata(myMultiAssay)
myMultiAssay <- MultiAssayExperiment(objlist, patient.data, dfmap)
myMultiAssay
```
