---
title: "Writing R Functions, 2018 Data Camp Course with Hadley Wickham"
author: "Jeffrey Long"
date: "10/11/2018"
output: html_document
---

`` `I dwell in Possibility —` ``

Emily Dickinson #466


## Writing R Functions

This R Markdown file follows the Data Camp course on R Functions team taught by Hadley Wickham, [https://campus.datacamp.com/courses/writing-functions-in-r/](https://campus.datacamp.com/courses/writing-functions-in-r/).

These are my notes.


```{r basicFunctionTemplate}
# setwd("/Users/jeffreylong/R/rFunctions/")
# setwd("/Users/jeffreylong/software/R/rFunctions")

# my_fun <- function(arg1, arg2) {
# 	body
# }

add <- function(x, y=1) {
	x + y
}

formals(add)

body(add)

environment(add)
```

The last expression evaluated in 
a function is the return value

```{r}
f <- function(x) {
	if (x < 0) {
	  -x
	} else {
	  x
	}
}

f(-5)
f(15)
```

Use return() to exit function and 
return value for special cases.

One can reassign a function name.
```{r}
mean2 <- mean
mean2(1:10)
```

### ANONYMOUS FUNCTIONS
Functions which don't need to be assigned 
```{r}
function(x) { x + 1 }

(function(x) { x + 1 })(2)
```
Anonymous functions must be all on one line.

Similar to creating a vector of random
 numbers and not saving it.

```{r}
# Define ratio() function
ratio <- function(x, y) {
  x / y
}
# Call ratio() with arguments 3 and 4
ratio(3,4)

ratio(x = 3, y = 4)
```

```{r}
?mean

# Usage: mean(x, trim = 0, na.rm = FALSE, …)
# Rewrite the call to follow best practices
mean(0.1,x=c(1:9, NA),TRUE)

mean(c(1:9, NA), trim = 0.1, na.rm = TRUE)
```

## Two types of vectors in R

Atomic vectors of six types: logical, integer, double, character, complex, and raw

Lists, a.k.a. recursive vectors, because lists can contain other lists

Atomic vectors are homogenous, lists can be heterogeneous

Its type, find with typeof()

```{r}
typeof(letters)
typeof(1:10)
```

Its length, find with:

$length()$

Check if value is NA:

$is.na()$

## Subsetting with [, [[ or $

[ extract  a sublist

[[ and $ extract elements, remove a level of hierarchy

```{r}

a <- list(
  element1 = 7:9,
  element2 = "Sylvan Woods",
  element3 = pi
)

str(a[1])

str(a[[1]])

a[[1]]

a[[1]][1]

a[[1]][[1]]

```


### Subsetting Practice with Tricky List
```{r}
df1 <- data.frame(
  a = 1:11,
  b = rnorm(11),
  c = rnorm(11),
  d = rnorm(11)
)

fit <- lm(a ~ b + c + d, data = df1)
tricky_list <- summary(fit)
tricky_list
## coefficients(fit) # model coefficients
## confint(fit, level=0.95) # CIs for model parameters 
## fitted(fit) # predicted values
## residuals(fit) # residuals
## anova(fit) # anova table 
## vcov(fit) # covariance matrix for model parameters 
## influence(fit) # regression diagnostics
## # diagnostic plots 
## layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page 
## plot(fit)
## # compare models
## fit1 <- lm(a ~ b + c + d, data = df1)
## fit2 <- lm(a ~ b + c, data = df1)
## anova(fit1, fit2)
## #install.packages("DAAG")
## library("DAAG")
## cv.lm(data = df1, fit, m=3) # 3 fold cross-validation
```

```{r}
# 2nd element in tricky_list 
typeof(tricky_list[[2]])

# Element called x in tricky_list
typeof(tricky_list[[3]])

# 2nd element inside the element called x in tricky_list
typeof(tricky_list[['x']][[2]])

names(tricky_list)
str(tricky_list)
```

```{r}
# Guess where the regression model is stored
names(tricky_list[['call']])

str(tricky_list[['call']])
# Subset the coefficients element
tricky_list[['coefficients']]
 
# Subset the wt element
tricky_list[['coefficients']][[2,3]]

```


##  for Loops in R

```{r}
primes_list <- list(2, 3, 5, 7, 11, 13, 17, 19, 23, 29, 31, 37, 41, 43)
for ( i in 1:length(primes_list)) { # sequence
	print(primes_list[[i]])              # body
}
```

```{r}
for (i in 1:ncol(df1)) {
	print(median(df1[[i]]))
}
```

### seq_along()

```{r}
# Replace the 1:ncol(df) sequence
for (i in seq_along(df1)) {
  print(median(df1[[i]]))
}

# Create an empty data frame
empty_df <- data.frame()

# Repeat for loop to verify there is no error
for (i in seq_along(empty_df)) {
  print(median(empty_df[[i]]))
}
```

### vector()

Has two arguments: the type of the vector ("logical", "integer", "double", "character", etc.) and the length of the vector.

```{r}
# Create new double vector: output
output <- vector("double", ncol(df1))
output

# Alter the loop
for (i in seq_along(df1)) {
  
# Change code to store result in output
output[[i]] = median(df1[[i]])
}
# Print output
output
```

```{r eval = FALSE}
# FUNCTIONS
df <-  df1
df$a <- (df$a - min(df$a, na.rm = TRUE)) /
  (max(df$a, na.rm = TRUE) - min(df$a, na.rm = TRUE))
df$b <- (df$b - min(df$b, na.rm = TRUE)) /
  (max(df$b, na.rm = TRUE) - min(df$b, na.rm = TRUE))
df$c <- (df$c - min(df$c, na.rm = TRUE)) /
  (max(df$c, na.rm = TRUE) - min(df$c, na.rm = TRUE))
df$d <- (df$d - min(df$d, na.rm = TRUE)) /
  (max(df$d, na.rm = TRUE) - min(df$d, na.rm = TRUE))

#########################################

# Better FUNCTIONS
df$a <- rescale01(df$a)
df$b <- rescale01(df$b)
df$c <- rescale01(df$c)
df$d <- rescale01(df$d)

#########################################

# BEST PURRR
install.packages("purrr")
library("purrr")
df[] <- map(df, rescale01)
```

```{r}
# Define example vector x
x <- 1:10
x

# Rewrite this snippet to refer to x
(x - min(x, na.rm = TRUE)) /
    (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))

# Define rng
rng = range(x, na.rm = TRUE)
# Rewrite this snippet to refer to the elements of rng
(x - min(x, na.rm = TRUE)) / (rng[[2]] - rng[[1]])

# Use the function template to create the rescale01 function
rescale01 <- function(x) {
  rng = range(x, na.rm = TRUE)
  (x - min(x, na.rm = TRUE)) / (rng[[2]] - rng[[1]])
  }

# Test your function, call rescale01 using the vector x as the argument
rescale01(x)
```

```{r}
# Define example vectors x and y
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3,  4)

# Count how many elements are missing in both x and y
sum(is.na(x) & is.na(y))

# Turn this snippet into a function: both_na()
both_na <- function(x, y) {
  sum(is.na(x) & is.na(y))
}

# Define x, y1 and y2
x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

# Removing last element of y2 
y3 <- y2[-length(y2)]
y3


# Call both_na on x, y1
both_na(x,y1)

# Call both_na on x, y2
both_na(x,y2)
?qnorm
```



```{r}
# Rewrite mean_ci to take arguments named level and x
mean_ci <- function(x, level = 0.95) {
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - level
  mean(x) + se * qnorm(c(alpha / 2, level / 2))
}
x <- c(1,2,3,4,5,6,7,8,9,10)
mean_ci(x)
mean_ci(x, .9)
mean_ci(x, .85)
mean_ci(x, .8)
```

```{r}
# Alter the mean_ci function
mean_ci <- function(x, level = 0.95) {
  if (length(x) == 0) {
    warning("`x` was empty", call. = FALSE)
    interval <- c(-Inf, Inf)
    return(interval)
  } else {
    se <- sd(x) / sqrt(length(x))
    alpha <- 1 - level
    interval <- mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
    return(interval)
  }
}
```

### New function definition

```{r}
f <- function(x, y) {
  x[is.na(x)] <- y
  cat(sum(is.na(x)), y, "\n")
  x
}

# Define a numeric vector x with the values 1, 2, NA, 4 and 5

x <- c(1,2,NA,4,5)

# Call f() with the arguments x = x and y = 3

f(x, y = 3)

# Call f() with the arguments x = x and y = 10

f(x, y = 10)

# New replace_missings funciton definition

replace_missings <- function(x, replacement) {
  # Define is_miss
  is_miss <- is.na(x)
  
  # Rewrite rest of function to refer to is_miss
  x[is_miss] <- replacement
  cat(sum(is_miss), replacement, "\n")
  x
}

x <- c(1,2,NA,4,5)
replace_missings(x, 88)


df <- data.frame(
  x = rnorm(10),
  y = rnorm(10),
  z = rnorm(10)
)
df$z <- c(-0.6264538, NA, -0.8356286, 1.5952808, 0.3295078, -0.8204684, 0.4874291, 0.7383247, 0.5757814, NA)

replace_missings <- function(x, replacement) {
  is_miss <- is.na(x)
  x[is_miss] <- replacement
  
# Rewrite to use message()
  message(sum(is_miss), " missing values replaced by the value ", replacement)
  x
}

# Check your new function by running on df$z
replace_missings(df$z, replacement = 0)
```


## Chapter 3 Data Camp R Functions Class by Hadley Wickam

```{r}
##########################################
# make the dataframe
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

median(df[[1]])
median(df[[2]])
median(df[[3]])
median(df[[4]])

##########################################

# Initialize output vector
output <- numeric(ncol(df))
# Fill in the body of the for loop
for (i in seq_along(df)) {            
  output[i] <- median(df[[i]])
}

# View the result
output

##########################################

# Turn this code into col_median()
col_median <- function(df) {
  output <- numeric(ncol(df))
  for (i in seq_along(df)) {            
    output[[i]] <- median(df[[i]])      
  }
  output
}

col_median(df)

##########################################

col_mean <- function(df) {
  output <- numeric(ncol(df))
  for (i in seq_along(df)) {
    output[[i]] <- mean(df[[i]])
  }
  output
}

##########################################

col_sd <- function(df) {
  output <- numeric(ncol(df))
  for (i in seq_along(df)) {
    output[[i]] <- sd(df[[i]])
  }
  output
}

##########################################

# Add a second argument called power
f <- function(x, power) {
    # Edit the body to return absolute deviations raised to power
    abs(x - mean(x))^power
}
#x <- c(4,5,6,7,8)
#f(x,2)

##########################################

col_summary <- function(df, fun) {
  output <- numeric(ncol(df))
  for (i in seq_along(df)) {
    output[[i]] <- fun(df[[i]])
  }
  output
}

##########################################

# Find the column medians using col_median() and col_summary()
col_median(df)
col_summary(df, median)

# Find the column means using col_mean() and col_summary()
col_mean(df)
col_summary(df, mean)

# Find the column IQRs using col_summary()
col_summary(df, IQR)

##########################################
```

## apply Family

Aside to Datacamp Intermediate R Class on lapply and sapply

<https://campus.datacamp.com/courses/intermediate-r/chapter-4-the-apply-family>

### lapply

```{r}
nyc <- list(pop = 8405837,
            boroughs = c("Manhattan", "Bronx", "Brooklyn", "Queens", "Staten Island"),
            capital = FALSE)

for (info in nyc) {
  print(class(info))
}
##################
lapply(nyc, class)

##########################################

cities <- c("New York", "Paris", "London", "Tokyo", "Rio de Janeiro")
num_chars <- c()
for(i in 1:length(cities)) {
  num_chars[i] <- nchar(cities[i])
}
num_chars
##################
lapply(cities, nchar)
##################
#Turn list into a vector
unlist(lapply(cities, nchar))

##########################################

oil_prices <- list(2.37, 2.49, 2.22, 2.47, 2.32)
triple <- function(x) {
  3 * x
}
result <- lapply(oil_prices, triple)
str(result)
unlist(result)
##################
multiply <- function(x, factor) {
  x * factor
}
times3 <- lapply(oil_prices, multiply, factor = 3)
unlist(times3)
times4 <- lapply(oil_prices, multiply, factor = 4)
unlist(times4)

##########################################
##########################################
##########################################


# The vector pioneers 
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")

# Split names from birth year
split_math <- strsplit(pioneers, split = ":")

# Convert to lowercase strings: split_low
split_low <- lapply(split_math, tolower)

# Take a look at the structure of split_low
str(split_low)

select_first <- function(x) {
  x[1]
}

# Apply select_first() over split_low: names
names <- lapply(split_low, select_first)

# Write function select_second()
select_second <- function(x) {
  x[2]
}

# Apply select_second() over split_low: years
years <- lapply(split_low, select_second)

##################
# Anonymous functions in lapply

# split_low has been created for you
split_low

# Transform: use anonymous function inside lapply
lapply(split_low, function(x) {x[1]})

# Transform: use anonymous function inside lapply
lapply(split_low, function(x) {x[2]})

##################
# Generic select function
select_el <- function(x, index) {
  x[index]
}

# Use lapply() twice on split_low: names and years
names <- lapply(split_low, select_el, index = 1)
years <- lapply(split_low, select_el, index = 2)

unlist(names)
unlist(years)

##################
# Apply function that return NULL
lapply(list(1, "a", TRUE), str)

str(TRUE)
```


### sapply

Simplify Apply.

Output is a named vector.

sapply calls lapply to apply function over each item of the vector. Then uses simplify to array function to convert the list that lapply created to an array.

<https://campus.datacamp.com/courses/intermediate-r/chapter-4-the-apply-family?ex=7>

```{r}
##########################################
##########################################
##########################################
# sapply
unlist(lapply(cities, nchar))
sapply(cities, nchar) 
sapply(cities, nchar, USE.NAMES = FALSE) 

# Returns the min and max letter of string in the alphabetical order
first_and_last <- function(name) {
  name <- gsub(" ", "", name)
  letters <- strsplit(name, split = "")[[1]]
  c(first = min(letters), last = max(letters))
}
first_and_last("New York")

##################

# Yields a matrix of first and last letters
sapply(cities, first_and_last)

# Unable to simplify?
unique_letters <- function(name) {
  name <- gsub(" ", "", name)
  letters <- strsplit(name, split = "")[[1]]
  unique(letters)
}
unique_letters("London")
lapply(cities, unique_letters)
sapply(cities, unique_letters)

```

sapply tries to simplify list to array. 

```{r}
##########################################
temp <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

# Use lapply() to find each day's minimum temperature
lapply(temp, min)

# Use sapply() to find each day's minimum temperature
sapply(temp, min)

# Use lapply() to find each day's maximum temperature
lapply(temp, max)

# Use sapply() to find each day's maximum temperature
sapply(temp, max)

##################
save(temp, file = "temp.RData")
##################

# Finish function definition of extremes_avg
extremes_avg <- function(x) {
  ( min(x) + max(x) ) / 2
}

# Apply extremes_avg() over temp using sapply()
sapply(temp, extremes_avg)

# Apply extremes_avg() over temp using lapply()
lapply(temp, extremes_avg)

##################

# Create a function that returns min and max of a vector: extremes
extremes <- function(x) {
  c(min = min(x), max = max(x))
}

# Apply extremes() over temp with sapply()
sapply(temp, extremes)

# Apply extremes() over temp with lapply()
lapply(temp, extremes)

##################

# Definition of below_zero()
below_zero <- function(x) {
  return(x[x < 0])
}

# Apply below_zero over temp using sapply(): freezing_s
freezing_s <- sapply(temp, below_zero)

# Apply below_zero over temp using lapply(): freezing_l
freezing_l <- lapply(temp, below_zero)


# Are freezing_s and freezing_l identical?
identical(freezing_s, freezing_l)

##################

# Definition of print_info()
print_info <- function(x) {
  cat("The average temperature is", mean(x), "\n")
}

# Apply print_info() over temp using sapply()
sapply(temp, print_info)

# Apply print_info() over temp using lapply()
lapply(temp, print_info)
```

### vapply

apply function over list or vector

$vapply(x, FUN, FUN.VALUE, ..., USE.NAMES = TRUE)$

explicitly specify output format

```{r}
##########################################
##########################################
##########################################
cities <- c("New York", "Paris", "London", "Tokyo", "Rio de Janeiro")
sapply(cities, nchar)
# vapply(x, FUN, FUN.VALUE, ..., USE.NAMES = TRUE)
vapply(cities, nchar, numeric(1))
first_and_last <- function(name) {
  name <- gsub(" ", "", name)
  letters <- strsplit(name, split = "")[[1]]
  c(first = min(letters), last = max(letters))
}
vapply(cities, first_and_last, character(2))
# These throw errors, so don't uncomment
## vapply(cities, first_and_last, character(1))
## vapply(cities, first_and_last, numeric(2))

##################
temp <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

# Definition of basics()
basics <- function(x) {
  c(min = min(x), mean = mean(x), max = max(x))
}

# Apply basics() over temp using vapply()
vapply(temp, basics, numeric(3))

##################

# Definition of the basics() function
basics <- function(x) {
  c(min = min(x), mean = mean(x), median = median(x), max = max(x))
}
vapply(temp, basics, numeric(4))

##################

# Convert to vapply() expression
# sapply(temp, max)
vapply(temp, max, numeric(1))

# Convert to vapply() expression
# sapply(temp, function(x, y) { mean(x) > y }, y = 5)
vapply(temp, function(x, y) { mean(x) > y }, y = 5, logical(1))
```

## purrr Intro

```{r}
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

# Load the purrr package
library(purrr)

# Use map_dbl() to find column means
map_dbl(df, mean)

# Use map_dbl() to column medians
map_dbl(df, median)

# Use map_dbl() to find column standard deviations
map_dbl(df, sd)

#########################################

planes <- data.frame(
  year = c(1956, 1975, 1977, 1996, 2010, NA),
  engines = c(4, 1, 2, 2, 2, 1),
  seats = c(102, 4, 139, 142, 20, 2),
  speed = c(232, 108, 432, NA, NA, NA)
)
planes

# Find the mean of each column
map_dbl(planes, mean)

# Find the mean of each column, excluding missing values
map_dbl(planes, mean, na.rm = TRUE)

# Find the 5th percentile of each column, excluding missing values
map_dbl(planes, quantile, probs = 0.05, na.rm = TRUE)

#########################################
df3 <- data.frame(
  A = rnorm(10),
  B = c("A","B","A","B","A","B","A","B","A","B"),
  C = c(1,2,3,4,5,6,7,8,9,10),
  D = rnorm(10)
)
# Find the columns that are numeric
map_lgl(df3, is.numeric)

# Find the type of each column
map_chr(df3, typeof)

# Find a summary of each column
map(df3, summary)

```
```{r}
# Just a quick aside to convert to uppercase
library(stringr)
stringr::str_to_upper("anonymous function")
```

### Specifying .f

Some shortcuts for anonymous function reference and subsetting

```{r}
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
# mapping an existing fuction of one you defined
map(df, summary)
# mapping an ANONYMOUS FUNCTION
map(df, function(x) sum(is.na(x)))
# mapping an anonymous function defined using a formula SHORTCUT
map(df, ~ sum(is.na(.)))

```

### Subsetting 

Shortcut for SUBSETTING

```{r}
list_of_results <- list(
  list(a = 1, b = "A"),
  list(a = 2, b = "B"),
  list(a = 3, b = "C")
)

# An Anonymous Function
map_dbl(list_of_results, function(x) x[["a"]])

# Shortcut for string subsetting
map_dbl(list_of_results, "a")

# Shortcut for integer subsetting
map_dbl(list_of_results,1)
```

## Purrr

Writing anonymous functions takes a lot of extra key strokes, so purrr provides a shortcut that allows you to write an anonymous function as a one-sided formula instead.

In R, a one-sided formula starts with a ~, followed by an R expression. In purrr's map functions, the R expression can refer to an element of the .x argument using the . character.

Let's take a look at an example. Imagine, instead of a regression on each data frame in cyl, we wanted to know the mean displacement for each data frame. One way to do this would be to use an anonymous function:

$map\_dbl(cyl, function(df) mean(df\$disp))$

To perform the same operation using the formula shortcut, we replace the function definition (function(df)) with the ~, then when we need to refer to the element of cyl the function operates on (in this case df), we use a ..

$map\_dbl(cyl, ~ mean(.\$disp))$

See how much less typing it involves! It also saves you from coming up with an argument name. Can you rewrite our previous anonymous function using this formula shortcut instead?

```{r}
# Split the data frame mtcars based on the unique values in the cyl column
cyl <- split(mtcars, mtcars$cyl)

# list of 3 dataframes, one for each 4, 6, and 8 cylinders
# Examine the structure of cyl
str(cyl)

# Extract the first element into four_cyls
four_cyls <- cyl[[1]]

# Fit a linear regression model of mpg on wt using four_cyls
lm(mpg ~ wt, data = four_cyls)

# Turn the snippet of code into a function
fit_reg <- function(df) {
  lm(mpg ~ wt, data = df)
}

# Pass the function into map()
map(cyl, fit_reg)

# Rewrite to call an anonymous function
map(cyl, function(df) lm(mpg ~ wt, data = df))

# Rewrite to use the formula shortcut instead
map(cyl, ~ lm(mpg ~ wt, data = .))
```

There are also some useful shortcuts that come in handy when you want to subset each element of the $.x$ argument. If the $.f$ argument to a map function is set equal to a string, let's say "name", then purrr extracts the "name" element from every element of $.x$.

This is a really common situation you find yourself in when you work with nested lists. For example, if we have a list of where every element contains an $a$ and $b$ element:

$$list_of_results <- list(\\
  list(a = 1, b = "A"), \\
  list(a = 2, b = "C"), \\
  list(a = 3, b = "D")\\
)$$

We might want to pull out the $a$ element from every entry. We could do it with the string shortcut like this:

$$map(list\_of\_results, "a")$$
Now take our list of regresssion models:

$$map(cyl, ~ lm(mpg ~ wt, data = .))$$

It might be nice to extract the slope coefficient from each model. You'll do this in a few steps: first fit the models, then get the coefficients from each model using the $coef()$ function, then pull out the $wt$ estimate using the string shortcut.

```{r}
# Save the result from the previous exercise to the variable models
models <- map(cyl, ~ lm(mpg ~ wt, data = .))
# Use map and coef to get the coefficients for each model: coefs
coefs <- map(models, coef)
# Use string shortcut to extract the wt coefficient 
map(coefs, "wt")
```

#### Using a numeric vector

Another useful shortcut for subsetting is to pass a numeric vector as the $.f$ argument. This works just like passing a string but subsets by index rather than name. For example, with your previous list_of_results:

$$ list_of_results <- list(\\
  list(a = 1, b = "A"), \\
  list(a = 2, b = "C"), \\
  list(a = 3, b = "D")\\
)$$
Another way to pull out the $a$ element from each list, is to pull out the first element:

$map(list_of_results, 1)$

Let's pull out the slopes from our models again, but this time using numeric subsetting. Also, since we are pulling out a single numeric value from each element, let's use $map_dbl()$.

```{r}
coefs <- map(models, coef)

# use map_dbl with the numeric shortcut to pull out the second element
map_dbl(coefs,2)
```

Putting it together with pipes

$purrr$ also includes a pipe operator: $\%>\%$. The pipe operator is another shortcut that saves typing, but also increases readability. The explanation of the pipe operator is quite simple: $x \%>\% f(y)$ is another way of writing $f(x, y)$. That is, the left hand side of the pipe, $x$, becomes the first argument to the function, $f()$, on the right hand side of the pipe.

Take a look at our code to get our list of models:

$$cyl <- split(mtcars, mtcars$cyl) \\
map(cyl, ~ lm(mpg ~ wt, data = .))$$

We split the data frame $mtcars$ and save it as the variable $cyl$. We then pass $cyl$ as the first argument to map to fit the models. We could rewrite this using the pipe operator as:

$$split(mtcars, mtcars$cyl) \%>\% 
  map(~ lm(mpg ~ wt, data = .))$$
  
We read this as "split the data frame $mtcars$ on $cyl$, then use $map()$ on the result."

One of the powerful things about the pipe is we can chain together many operations. Here is our complete code, written with pipes, instead assigning each step to a variable and using it in the next step:

$$mtcars \%>\% \\
  split(mtcars\$cyl) \%>\% \\
  map(~ lm(mpg ~ wt, data = .)) \%>\% \\
  map(coef) \%>\% \\
  map_dbl("wt")$$
  
We've written some code in the editor to pull out the $R^2$ from each model. Rewrite the last two lines to use a pipe instead.

```{r}
# Define models (don't change)
models <- mtcars %>% 
  split(mtcars$cyl) %>%
  map(~ lm(mpg ~ wt, data = .))

# Rewrite to be a single command using pipes 
# summaries <- map(models, summary)
# map_dbl(summaries, "r.squared")
summaries <- models %>%
  map(summary) %>%
  map_dbl("r.squared")
```

## Chapter 4 

### Dealing with Failure

safely() takes a function and returns a function.

Other adverbs for unusual output include 

safely(), captures the successful result or the error, always returns a list;

possibly(), always succeeds, you give it a default value to return when there is an error;

quietly(), captures printed output, messages, and warnings instead of capturing errors.

```{r}
library(purrr)
# log_list <- map(long_list, log)
# Error
log_list <- map(long_list, safely(log))
str(log_list)

# safely returns $result and $error
safe_log <- safely(log)
safe_log(10)
safe_log("a")

```

Creating a safe function

$safely()$ is an adverb; it takes a verb and modifies it. That is, it takes a function as an argument and it returns a function as its output. The function that is returned is modified so it never throws an error (and never stops the rest of your computation!).

Instead, it always returns a list with two elements:

$result$ is the original result. If there was an error, this will be $NULL$.
$error$ is an error object. If the operation was successful this will be $NULL$.

Let's try to make the $readLines()$ function safe.

```{r}
# Create safe_readLines() by passing readLines() to safely()
safe_readLines <- safely(readLines)

# Call safe_readLines() on "http://example.org"
example_lines <- safe_readLines("http://example.org")
example_lines

# Call safe_readLines() on "http://asdfasdasdkfjlda"
nonsense_lines <- safe_readLines("http://asdfasdasdkfjlda")
nonsense_lines
```

Using map safely

One feature of $safely()$ is that it plays nicely with the $map()$ functions. Consider this list containing the two URLs from the last exercise, plus one additional URL to make things more interesting:

$$urls <- list(\\
  example = "http://example.org",\\
  rproj = "http://www.r-project.org",\\
  asdf = "http://asdfasdasdkfjlda"\\
)$$
We are interested in quickly downloading the HTML files at each URL. You might try:

$$map(urls, readLines)$$
But it results in an error, $Error in file(con, "r") : cannot open the connection$, and no output for any of the URLs. Go on, try it!

We can solve this problem by using our safe_readLines() instead.