---
title: "Unit Testing"
author: "Jeffrey Long"
date: "9/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Libraries

```{r}
# setwd("~/software/R/unitTesting")
install.packages("testthat")
library(testthat)
```


## Resources

[Unit Testing in R](https://towardsdatascience.com/unit-testing-in-r-68ab9cc8d211)

[Jim Hester's lintr package](https://github.com/jimhester/lintr)

[Intro to Unit Testing in R](https://katherinemwood.github.io/post/testthat/)

[testthat Reference](https://testthat.r-lib.org/reference/index.html)

[testthat Package Doc](https://cran.r-project.org/web/packages/testthat/testthat.pdf)


[http://r-pkgs.had.co.nz/tests.html](http://r-pkgs.had.co.nz/tests.html)


### From Unit Testing in R

```{bash}

cat > "~/software/R/unitTesting/increment.R"
# Example for simple function to be tested
#
# 1. Change working directory
# 2. Call: testthat::test_dir('tests') within RStudio or R

increment <- function(value) {
  value + 1
}

mkdir tests

cat > "~/software/R/unitTestingtests/tests/test_increment.R"
source("../increment.R", chdir = TRUE)   # NOT USED FOR PACKAGES
library(testthat)

test_that("single number", {
  expect_equal(increment(-1), 0)
  expect_equal(increment(0), 1)
})

test_that("vectors", {
  expect_equal(increment(c(0,1)), c(1,2))
})

test_that("empty vector", {
  expect_equal(increment(c()), c())
})

test_that("test NA", {
  expect_true(is.na(increment(NA)))
})
```

#### Running Tests

```{r}

testthat::test_dir("tests")

```


Structure of test declaration function is

$testthat::test\_that(name, expression)$

 * The first argument assigns the test a name for identifying it. 
 
 * The second argument is an R expression which shall use the expect_* assertions. 
 
 Whenever an assertion does not hold the test is aborted and marked as failed.


### From Intro to Unit Testing in R

[https://katherinemwood.github.io/post/testthat/](https://katherinemwood.github.io/post/testthat/)

[https://github.com/katherinemwood/unit_testing](https://github.com/katherinemwood/unit_testing)


```{r}
print(testing_data)

# Add context for testing
context('testing data integrity') 

# Check number of cols in our data equal the number of variables
expect_equal(ncol(testing_data), 2)    # this one passes
expect_equal(ncol(testing_data), 3)    # this one fails

# Test the number of coefficients is equal
expect_equal(length(test.mod$coefficients), 2)    # this one passes

# Check for missing values
expect_identical(testing_data, na.omit(testing_data))   # this one passes

# expect_equivalent ignores attributes of comparators

expect_that(testing_data, is_a('data.frame'))
# same as
expect_is(testing_data,'data.frame')

# test  factor levels in data are in model
expect_equivalent(levels(model_data$x), unlist(test.mod$xlevels))

# test mean of group 1 equals intercept
expect_equivalent(mean(model_data$y[model_data$x == 'c1']), test.mod$coefficients['(Intercept)'])

# test printed values
expect_that(even_odd(1), prints_text('odd'))     # this one passes
expect_that(even_odd(2), prints_text('odd'))    # this one fails

# group tests in a test file
test_file('tests/test_dummy.R')


```
